\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation

### 2018-05-04 ccs/ex.tex
yao comparison

\figureref{toolchain} shows an overview of the \tool
toolchain. We give a brief overview of each of these phases below.
\vspace{-0.15in}
\paragraph{Source language.}
Consider the example $w^Tx >b$ from \sectionref{intro}, where
$w$ and $b$ are the server's input (a classifier) and $x$ is
the client's input vector. \figureref{ex-sml} shows the \tool code
for this example. The code first reads the server's (resp. client's) input using
 \ls{input1} (resp. \ls{input2}). It then uses a
\ls{for} loop to compute \ls{acc}, the inner product of \ls{w} and
\ls{x}. Finally, the code compares \ls{acc}
with \ls{b} and outputs this only to the client using \ls{output2}.

\tool source language is a simple, imperative language that enables
the programmers to express \mpc computations in terms of their
``ideal'' functionalities, without dealing with any cryptographic
details. The language provides multi-dimensional arrays, conditional
expressions (the ternary $?\::$ operator), \ls{for}~loops,
\ls{if}~statements, and special syntax for input/output from each party.
\vspace{-0.1in}
\paragraph{\tool compiler.}
\tool compiler takes as input a source program and produces a C++
program as output. \figureref{ex-aby} shows the output code for
the example in \figureref{ex-sml} -- this is also how a program written directly in
  ABY would look like. The output program
contains party-specific code for inputs and outputs
(\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
computation.

@@

\paragraph{Circuit generation and evaluation.} The next step in
\tool is to
compile the output C++ code and execute it. Doing so evaluates away the
public parts of the program, including the array accesses, and
generates a  circuit comprising of arithmetic and boolean gates,
with appropriate conversion gates. The circuit is then evaluated using
a \mpc protocol.
\vspace{-0.1in}
\paragraph{Advantages of \tool.} We can now concretely see the advantages of \tool. Unarguably, it is
easier for a developer to program and get the code right in
\figureref{ex-sml}, rather than the code in
\figureref{ex-aby}.
% (as we mentioned in Section~\ref{sec:intro},
%other high-level languages are limited in their use of a single,
%arithmetic or boolean, crypto back-end). 
\tool also enables the
programmer to easily modify their code, while the compiler takes care
of efficiency. For example, consider in \figureref{ex-sml} a
change from multiplication to bitwise-or in the \ls{for} loop. It
turns out that in this case, it is more efficient to do both
the addition and bitwise-or using boolean circuits (if the addition is
done using arithmetic, the conversion cost starts to take
over). In \tool, the programmer simply needs to change one operator in
the source code, and the compiler generates efficient code that uses
boolean addition. Whereas, if the programmer was writing ABY
code, she either has to sacrifice performance, or would have to
revisit many parts of the circuit and change them. In summary, \tool raises the level of abstraction for the
programmer, and generates efficient \mpc protocols automatically,
while its metatheory provides strong correctness and security
guarantees.

>>>>

\figureref{toolchain} shows an overview of the \tool
toolchain. We give a brief overview of each of these phases below.

\paragraph{Source language.}
Consider the example $w^Tx >b$ from \sectionref{intro}, where
$w$ and $b$ are the server's input (a classifier) and $x$ is
the client's input vector. \figureref{ex-sml} shows the \tool code
for this example. The code first reads the server's (resp. client's) input using
 \ls{input1} (resp. \ls{input2}). It then uses a
\ls{for} loop to compute \ls{acc}, the inner product of \ls{w} and
\ls{x}. Finally, the code compares \ls{acc}
with \ls{b} and outputs this only to the client using \ls{output2}.

\tool source language is a simple, imperative language that enables
the programmers to express \mpc computations in terms of their
``ideal'' functionalities, without dealing with any cryptographic
details. The language provides multi-dimensional arrays, conditional
expressions (the ternary $?\::$ operator), \ls{for}~loops,
\ls{if}~statements, and special syntax for input/output from each party.

\paragraph{\tool compiler.}
\tool compiler takes as input a source program and produces a C++
program as output. \figureref{ex-aby} shows the output code for
the example in \figureref{ex-sml} -- this is also how a program written directly in
  ABY would look like. The output program
contains party-specific code for inputs and outputs
(\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
computation.

@@

\paragraph{Circuit generation and evaluation.} The next step in
\tool is to
compile the output C++ code and execute it. Doing so evaluates away the
public parts of the program, including the array accesses, and
generates a  circuit comprising of arithmetic and boolean gates,
with appropriate conversion gates. The circuit is then evaluated using
a \mpc protocol.

\paragraph{Advantages of \tool} We can now concretely see the advantages of \tool. Unarguably, it is
easier for a developer to program and get the code right in
\figureref{ex-sml}, rather than the code in
\figureref{ex-aby}.
% (as we mentioned in Section~\ref{sec:intro},
%other high-level languages are limited in their use of a single,
%arithmetic or boolean, crypto back-end). 
\tool also enables the
programmer to easily modify their code, while the compiler takes care
of efficiency. For example, consider in \figureref{ex-sml} a
change from multiplication to bitwise-or in the \ls{for} loop. It
turns out that in this case, it is more efficient to do both
the addition and bitwise-or using boolean circuits (if the addition is
done using arithmetic, the conversion cost starts to take
over). In \tool, the programmer simply needs to change one operator in
the source code, and the compiler generates efficient code that uses
boolean addition. Whereas, if the programmer was writing ABY
code, she either has to sacrifice performance, or would have to
revisit many parts of the circuit and change them. In summary, \tool raises the level of abstraction for the
programmer, and generates efficient \mpc protocols automatically,
while its metatheory provides strong correctness and security
guarantees.

\subsection{Comparison with Yao-only approach}
Before we provide the formalization of our compiler, we show that for performance it it critical to use a mix of arithmetic and boolean circuits.
We demonstrate the performance benefits of mixed computation over boolean-only compute by comparing with previous tools CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm} that are based on Yao's garbled circuits. As we already mentioned that we are not the first to observe that Yao-only approach does not scale well beyond examples with small numbers of multiplications \cite{aby,tasty,autos,secureml,minionn,chameleon}, especially for large-scale multiplications commonly found in Machine Learning (ML) algorithms -- the main use case of \tool. Indeed, this is one of the main drivers behind development of the specialized MPC protocols in previous works \cite{secureml,minionn,shafindss} with which we compare the performance of our automatically generated protocols in \sectionref{eval}.

The largest example benchmark in CBMC-GC is multiplication of two $8\times 8$ matrices for which it generates <FILL> gates and takes <FILL> seconds to run. In contrast, in \tool we use arithmetic sharing for this compute that generates <FILL> gates and completes in <FILL> seconds. We tried running CBMC-GC for multiplying two $20\times 20$ matrices and it timed out after 5 hours. In contrast, most of our benchmarks are much bigger. For example, we run a moderate size DNN for <FILL> in <FILL> seconds which contains a multiplication of <FILL> matrix and <FILL> matrix as a sub-computation. On the other hand, ObliVM scales to larger benchmarks with its use of Yao-pipelining \cite{yao-pipe}. We ran inner-product computations of varying lengths followed by a comparison in both ObliVM and \tool (see Table~\cite{oblivm-compare}). Our experiments show that we are atleast 30x faster than ObliVM due to our use of arithmetic sharing for inner-product computation.

### 2018-05-04 ccs/intro.tex
starting ccs repo
yao comparison

 Today it is hard for developers to program secure applications
 using cryptographic techniques. Typical developers lack a deep
 understanding of cryptographic protocols, and cannot be
 expected to use them correctly and efficiently on their own.
 Ideally, a developer would declare the functionality in a, general
 purpose, high-level programming language and a
 tool, e.g., a compiler would generate an efficient protocol that
 implements the functionality securely, while hiding the cryptography
 behind-the-scenes.

@@

She can program the computation in one of the several programmer
friendly, domain-specific languages (such
as Fairplay~\cite{fairplay}, Wysteria~\cite{wysteria},
ObliVM~\cite{oblivm}, CBMC-GC~\cite{cbmcgc}, SMCL~\cite{smcl},
Sharemind~\cite{sharemind}, \cite{lambdaps} etc.) that would
automatically compile it to a \mpc protocol. However, all of these
frameworks use cryptographic backends that take as input
the computation expressed either as a boolean
circuit~(\cite{yao,gmw}) or as an arithmetic circuit~(\cite{sss,viff,gentry}). 
%
The efficiency of the generated \mpc protocol is thus bounded by the
efficiency of representing the computation in \emph{one} of
these representations. For instance, multiplication of two
$\ell$-bit integers can either be expressed as a boolean circuit of
size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
gate.
%
For better efficiency, Alice would ideally like to compute
$w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
a boolean circuit.
%
Unfortunately, none of the above frameworks support combinations of
arithmetic and boolean circuits, and using different tools for
different parts of the computation is cumbersome and error-prone.

@@

\begin{figure}
\begin{lstlisting}[language=C,mathescape=true]
$\mathtt{
uint\;w[30] = input1();\;\;uint\;b = input1();
}$
$\mathtt{
uint\;x[30] = input2();
}$
$\mathtt{
uint\;acc = 0;
}$
$\mathtt{
for\;i\;in\;[0:30]\;\{\;acc\;=\;acc\;+\;(w[i]\;\times\;x[i]);\;\}
}$
$\mathtt{
output2((acc > b)\;?\;1\;:\;0)\;\textbf{{\color{blue}//only to party 2}}
}$
\end{lstlisting}
\caption{\tool code for $w^Tx >b$}
\label{fig:ex-sml}
\end{figure}
\vspace{-0.1in}
\paragraph{Ease of programming.} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, \figureref{ex-sml}
shows an \tool source
program for $w^Tx >b$. The program is quite similar to what a programmer might
write in C++ or Java. The simplicity of the
language comes with the usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, developers
don't bear the burden of getting cryptographic details right, 
code optimizations can be left to the compiler,
and it
is easy to maintain and modify the programs. Needless to say,
frameworks that expose low-level circuit APIs to the programmer do not
enjoy these benefits.
\vspace{-0.1in}
\paragraph{Cryptographic-cost aware compiler.} The \tool compiler
compiles a source program to a hybrid computation consisting
of \emph{public} and \emph{secret} parts. In the example above, for
instance, \tool compiler realizes that the array index~\ls{i} is
public, and generates non-cryptographic code for the array accesses.
Further, within the secret parts, \tool compiler is
aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators. Based on these costs, the compiler
automatically picks arithmetic or boolean
representations for different sub-parts, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an efficient \mpc protocol combining arithmetic
and boolean circuits, while the programmer remains
oblivious of all these cryptographic details. Indeed, \tool is the
first such cryptographic-cost aware compiler.
\vspace{-0.1in}
\paragraph{Scalability (secure code partitioning).} \mpc tools
often do not scale to large functionalities. The reason is that
most \mpc implementations use a circuit-like representation as an
intermediate language. Hence, the largest compute that can be done
securely is upper-bounded by the largest circuit that can fit in the
machine memory\footnote{Using swap and
disk space is feasible but it causes huge slowdown.}. This is a
show-stopper for applications like secure machine learning, secure
prediction, etc. that operate on large amounts of data.
%\aseem{Can we
%cite something here to support this claim?}
\tool addresses the scalability concern using a novel technique that
we call secure code partitioning (or partitioning in short). At
a high level, we decompose the original program into a sequence of small
sub-programs, which are then sequentially processed by \tool, while
appropriately threading the intermediate outputs
along. While this
addresses the scalability concern (the circuit
sizes of the sub-programs are now small enough to fit in the memory),
we still have to address
the security risk of revealing the intermediate outputs. \tool comes
to the rescue; it automatically inserts the required instrumentation
to ensure security of these intermediate outputs (\sectionref{pipe}). As
we show in our
evaluation, partitioning allows us to program large applications
in \tool.

%% We provide details of partitioning and prove the
%% correctness and security of this transformation
%% in \sectionref{pipe}
\vspace{-0.1in}
\paragraph{Formal guarantees.} We prove formal correctness and
security theorems for our compiler. The correctness theorem relates the
``trusted third party'' semantics of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. The theorem
guarantees that for all well-typed source programs, the two semantics
successfully terminate
(e.g., there are no array index out-of-bounds errors) with identical
observable outputs. For
the security theorem, we formally reduce the security
of our scheme against semi-honest (or, ``honest but curious") adversaries
 to the semi honest security of the  \mpc back-end. 
 We also prove a formal security theorem against semi-honest  adversaries for our partitioning
scheme (\sectionref{ld} and~\sectionref{pipe}).
%\vspace{-0.2in}
\paragraph{Evaluation.} We have implemented \tool
using ABY~\cite{aby} as the
cryptographic back-end. We evaluate \tool by implementing a wide range
of secure prediction benchmarks including linear and na\"{i}ve bayes
classifiers, decision trees, deep neural networks, state-of-the-art
classifiers from Tensorflow~\cite{tensorflow}
and \bonsai~\cite{bonsai}, and also the matrix factorization example
from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
three key points. First, \tool makes it convenient for general
programmers to write \mpc protocols. E.g. we provide the first \mpc
implementation of \bonsai~\cite{bonsai}, and it was programmed in the
high-level \tool source language by a non-cryptographer. Second, the
performance
of the protocols
generated by \tool are comparable to or better than (up to 19x) their state-of-the-art, hand-crafted implementations. Finally, we
demonstrate the usefulness of partitioning by implementing an application
that requires more than 300 million gates (\sectionref{impl} and~\sectionref{eval}).
\vspace{-0.1in}
\paragraph{Related Work.}
Before ABY~\cite{aby}, several works have proposed
combining secure computation protocols based on homomorphic
encryption and Yao's garbled circuits
(e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
and some have also developed tools that allow writing such
combinations (e.g. \cite{bogdanov,lone,tasty,autoS}). However, as Demmler et
al.~\cite{aby} observe, due
 to the high conversion cost between
homomorphic encryption and Yao's garbled circuits, these combined
protocols do not gain much performance over a single
protocol. 
Additionally, these prior works provide informal languages
or libraries that lack formal semantics and static guarantees.
%Can be added to related work
% Kerschbaum et al.~\cite{autoS}
%provide a scheme to automatically assign homomorphic encryption or
%garbled circuits to each operators in a computation that has been expressed as a sequence of dyadic operations.
%They conjecture that the problem is NP hard and gave a linear programming based solution
%as well as a quadratic time greedy heuristic. \tool  uses a simpler constant time heuristic
%to assign arithmetic secret sharing or garbled circuits to sub-computations
%in a richer language with control flow statements such as loops and branches.
 We provide a detailed survey of related
work in \sectionref{related}.

>>>>

 Today it is hard for developers to program secure applications
 using cryptographic techniques. Typical developers lack a deep
 understanding of cryptographic protocols, and cannot be
 expected to use them correctly and efficiently on their own.
 Ideally, a developer would declare the functionality in a general
 purpose, high-level programming language and a
 tool, e.g., a compiler would generate an efficient protocol that
 implements the functionality securely, while hiding the cryptography
 behind-the-scenes.

@@

She can program the computation in one of the several programmer
friendly, domain-specific languages (such
as Fairplay~\cite{fairplay}, Wysteria~\cite{wysteria},
ObliVM~\cite{oblivm}, CBMC-GC~\cite{cbmcgc}, SMCL~\cite{smcl},
Sharemind~\cite{sharemind}, \cite{lambdaps} etc.) that would
automatically compile it to a \mpc protocol. However, all of these
frameworks use cryptographic backends that take as input
the computation expressed either as a boolean
circuit~(\cite{yao,gmw}) or as an arithmetic circuit~(\cite{sss,viff,gentry}). 
%
The efficiency of the generated \mpc protocol is thus bounded by the
efficiency of representing the computation in \emph{one} of
these representations. For instance, multiplication of two
$\ell$-bit integers can either be expressed as a boolean circuit of
size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
gate.
\dg{It is well-known that boolean circuits are not suitable for doing arithmetic operations such as integer multiplications but are unavoidable for boolean operations such as comparison \cite{aby,tasty,autos,secureml,minionn,chameleon}.}
%
For better efficiency, Alice would ideally like to compute
$w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
a boolean circuit.
%

Unfortunately, none of the above frameworks support combinations of
arithmetic and boolean circuits, and using different tools for
different parts of the computation is cumbersome and error-prone.

@@

\begin{figure}
\begin{lstlisting}[language=C,mathescape=true]
$\mathtt{
uint\;w[30] = input1();\;\;uint\;b = input1();
}$
$\mathtt{
uint\;x[30] = input2();
}$
$\mathtt{
uint\;acc = 0;
}$
$\mathtt{
for\;i\;in\;[0:30]\;\{\;acc\;=\;acc\;+\;(w[i]\;\times\;x[i]);\;\}
}$
$\mathtt{
output2((acc > b)\;?\;1\;:\;0)\;\textbf{{\color{blue}//only to party 2}}
}$
\end{lstlisting}
\caption{\tool code for $w^Tx >b$}
\label{fig:ex-sml}
\end{figure}

\paragraph{Ease of programming.} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, \figureref{ex-sml}
shows an \tool source
program for $w^Tx >b$. The program is quite similar to what a programmer might
write in C++ or Java. The simplicity of the
language comes with the usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, developers
don't bear the burden of getting cryptographic details right, 
code optimizations can be left to the compiler,
and it
is easy to maintain and modify the programs. Needless to say,
frameworks that expose low-level circuit APIs to the programmer do not
enjoy these benefits.

\paragraph{Cryptographic-cost aware compiler.} The \tool compiler
compiles a source program to a hybrid computation consisting
of \emph{public} and \emph{secret} parts. In the example above, for
instance, \tool compiler realizes that the array index~\ls{i} is
public, and generates non-cryptographic code for the array accesses.
Further, within the secret parts, \tool compiler is
aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators. Based on these costs, the compiler
automatically picks arithmetic or boolean
representations for different sub-parts, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an efficient \mpc protocol combining arithmetic
and boolean circuits, while the programmer remains
oblivious of all these cryptographic details. Indeed, \tool is the
first such cryptographic-cost aware compiler.

\paragraph{Scalability (secure code partitioning).} \mpc tools
often do not scale to large functionalities. The reason is that
most \mpc implementations use a circuit-like representation as an
intermediate language. Hence, the largest compute that can be done
securely is upper-bounded by the largest circuit that can fit in the
machine memory\footnote{Using swap and
disk space is feasible but it causes huge slowdown.}. This is a
show-stopper for applications like secure machine learning, secure
prediction, etc. that operate on large amounts of data.
%\aseem{Can we
%cite something here to support this claim?}
\tool addresses the scalability concern using a novel technique that
we call secure code partitioning (or partitioning in short). At
a high level, we decompose the original program into a sequence of small
sub-programs, which are then sequentially processed by \tool, while
appropriately threading the intermediate outputs
along. While this
addresses the scalability concern (the circuit
sizes of the sub-programs are now small enough to fit in the memory),
we still have to address
the security risk of revealing the intermediate outputs. \tool comes
to the rescue; it automatically inserts the required instrumentation
to ensure security of these intermediate outputs (\sectionref{pipe}). As
we show in our
evaluation, partitioning allows us to program large applications
in \tool.

%% We provide details of partitioning and prove the
%% correctness and security of this transformation
%% in \sectionref{pipe}

\paragraph{Formal guarantees.} We prove formal correctness and
security theorems for our compiler. The correctness theorem relates the
``trusted third party'' semantics of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. The theorem
guarantees that for all well-typed source programs, the two semantics
successfully terminate
(e.g., there are no array index out-of-bounds errors) with identical
observable outputs. For
the security theorem, we formally reduce the security
of our scheme against semi-honest (or, ``honest but curious")
adversaries to the semi honest security of the  \mpc back-end. The
theorem provides protection against side-channels arising from
conditionals and memory access patterns.
 We also prove a formal security theorem against semi-honest
 adversaries for our partitioning
scheme (\sectionref{ld} and~\sectionref{pipe}).

\paragraph{Evaluation.} We have implemented \tool
using ABY~\cite{aby} as the
cryptographic back-end. We evaluate \tool by implementing a wide range
of secure prediction benchmarks including linear and na\"{i}ve bayes
classifiers, decision trees, deep neural networks, state-of-the-art
classifiers from Tensorflow~\cite{tensorflow}
and \bonsai~\cite{bonsai}, and also the matrix factorization example
from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
three key points. First, \tool makes it convenient for general
programmers to write \mpc protocols. E.g. we provide the first \mpc
implementation of \bonsai~\cite{bonsai}, and it was programmed in the
high-level \tool source language by a non-cryptographer. Second, the
performance
of the protocols
generated by \tool are comparable to or better than (up to 19x) their state-of-the-art, hand-crafted implementations. Finally, we
demonstrate the usefulness of partitioning by implementing an application
that requires more than 300 million gates (\sectionref{impl} and~\sectionref{eval}).

\paragraph{Related Work.}
Before ABY~\cite{aby}, several works have proposed
combining secure computation protocols based on homomorphic
encryption and Yao's garbled circuits
(e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
and some have also developed tools that allow writing such
combinations (e.g. \cite{bogdanov,lone,tasty,autoS}). However, as Demmler et
al.~\cite{aby} observe, due
 to the high conversion cost between
homomorphic encryption and Yao's garbled circuits, these combined
protocols do not gain much performance over a single
protocol. 
Additionally, these prior works provide informal languages
or libraries that lack formal semantics and static guarantees.
%Can be added to related work
% Kerschbaum et al.~\cite{autoS}
%provide a scheme to automatically assign homomorphic encryption or
%garbled circuits to each operators in a computation that has been expressed as a sequence of dyadic operations.
%They conjecture that the problem is NP hard and gave a linear programming based solution
%as well as a quadratic time greedy heuristic. \tool  uses a simpler constant time heuristic
%to assign arithmetic secret sharing or garbled circuits to sub-computations
%in a richer language with control flow statements such as loops and branches.
 We provide a detailed survey of related
work in \sectionref{related}.
